Bayesian exponential random graph modeling of KWL Burial network

### 2020-11-14 analysis/paper/paper.Rmd
add citation, inline code, and reproducibility section

<!-- Network construction and specification -->
We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language <!-- cite R --> with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008]. Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. Based on our hypothesis, we model a network with increased social inequality that is represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects. We modeled four node attributes for homophily effects, including age, gender, ritual activity, and the degree of wealth, to explore the correlation between the tendency of connected nodes to form relationships and the network ties overall. We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006]. Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others. This would be indicated by high popularity or degree values in the network statistics. According to our hypothesis, the burial evidence from after the European arrival will have a better fit with our model compared with the burial evidence from before European contact.

```{r}
burn_in <- 200
...
```

<!--Model fitting and prior setting--> 
After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations was 200 `r burn_in` and the number of iterations after the burn-in was 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. We use the normal distribution for the priors, as is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].

@@

>>>>

<!-- Network construction and specification -->
We used Bayesian inference on exponential random graph models (ERGMs) to quantify the relations among burials and test our hypothesis of social change using the R programming language [@Rlanguage2019] with the bergm package [@Caimo2014], which is built upon the statnet package for model specification and simulation procedures [@Hunter2008; @Morris2008; @Handcock2008]. Table \@ref(tab:ergm-terms-table) lists the configurations we used for the model specification for burials with the corresponding archaeological evidence. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. Based on our hypothesis, we model a network with increased social inequality that is represented by low reciprocity, low transitivity, high popularity, high age homophily, and less proximity structure effects. We modeled four node attributes for homophily effects, including age, gender, ritual activity, and the degree of wealth, to explore the correlation between the tendency of connected nodes to form relationships and the network ties overall. We also take into account the physical distance between burials as an indicator of a kinship-based relation since the deceased from the same family tend to be buried nearby [@LiandWu2006]. Our model could reveal the emergence of social inequality via the presence of a few individuals as network centers, having more relations with others. This would be indicated by high popularity or degree values in the network statistics. According to our hypothesis, the burial evidence from after the European arrival will have a better fit with our model compared with the burial evidence from before European contact.

```{r}
burn_in <- 200
main_iter <- 1000
aux_iter <- 10000
...
```

<!--Model fitting and prior setting--> 
After we set our model parameters, we simulated networks in a Bayesian framework using a Markov chain Monte Carlo (MCMC) algorithm. MCMC algorithms allow estimation of posterior distributions through direct random sampling the posterior without assuming the prior comes from any specific distribution [@Hamra2013]. We can obtain a posterior distribution by constructing a Markov chain that describes a sequence of moves from current state to the next state following probabilistic rules based on an algorithm. This enables a random or stochastic simulation in a long run where each move does not depend on the previous move. More chains ensures a more desirable posterior distribution that is close to the target distribution under study, or convergence. In Bayesian ERGMs, MCMC first selects a set of edges (or a set of empty pair of actors) with equal probability, and then switches to a pair of actors at random within the chosen set [@Caimo2011]. In our case, we set the number of chains to six. For each chain, the number of burn-in iterations was `r burn_in` and the number of iterations after the burn-in was `r main_iter`. We set the number of iterations used to simulate a network y′ at each iteration to `r aux_iter`. We use the normal distribution for the priors, as is typical in network analysis studies that assume networks to have low density and high transitivity, as are commonly found in the real world [@Caimo2017].

@@

# Reproducibility and open source materials

The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at xxx to enable re-use of materials and improve reproducibility and transparency [@Marwick2017]. Also in this version-controlled compendium [@Marwick2018] are the raw data for all the visualizations and tests reported here. All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository. The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.
